Citation Analysis of Computer Science Publications

### 2019-12-08 Scientometrics/comp.tex
Minor language / punctuation edits

Other classification systems are available such as the All Science Journal Classification (ASJC) developed and maintained by Scopus. the Web of Science research and categorical classifications from Clarivate Analytics, and the NSF classification system~\cite{nsf_classification}. Of these, Scopus and the Web of Science also include global bibliographies with citation links and proprietary unique identifiers for publications that can be matched, in some cases, to digital object identifiers (DOIs). All three rely on applying one or more journal-derived labels to articles. The inheritance, by articles, of labels meant to describe journals can be problematic in fields that evolve faster than their journal classifications. Shu and colleagues recently noted in a comparative study of the Chinese Science Citation Database (CSCD) and the Web of Science, that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. Others have also discussed and critiqued disciplinary assignments using journal based classifications~\cite{wang_large-scale_2016} \textcolor{blue}{need comprehensive list of citations}. Article classification systems have been constructed, however, which escape some of the criticisms of journal-based classification~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}.

@@

\subsection{Data}
 A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the DBLP dataset were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceedings. References cited by these publications were then extracted from Scopus (7,129,006), resulting in a total of 8,000,411 nodes (publications and references) and  44,296,381 edges that represent citations within the dataset. This dataset is referred to as \emph{comp} (see Fig~\ref{fig:ar_cp_annotation}).

@@

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus, which requires the number of clusters to be formed as an input parameter. In preliminary experiments, we varied the number of clusters to be formed between 10 and 50 clusters (data not shown).  At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}.  Consequently, we generated 20 clusters, labeled 0-19, using Graclus (Fig~\ref{fig:heatmap}).
We used conductance, as defined in Shun et al.~\cite{shun_parallel_2016}, to  evaluate clustering (smaller is better), noting that
conductance has been found to be ``the stand-alone quality metric that best indicates performance on the information recovery metrics" \cite{emmons2016analysis}.
% The quality of clustering was evaluated by measuring conductance as defined in Shun et al.~\cite{shun_parallel_2016}. 
%George, Tandy - add text about why conductance is good, and that small values are better
The data in \emph{comp} were clustered into 18, 20, or 22 clusters and in each case the highest numbered cluster (17, 19, or 21) had the greatest conductance value and the smallest number of nodes, suggesting that it effectively served as a container for `left over publications' during the Graclus process. Thus, we did not consider cluster 19 (the 20th cluster) when interpreting our results.  The remaining 19 clusters had conductance values ranging from 0.09 to 0.25 with a median conductance of 0.15.

@@

\paragraph{Clustering by Co-Citation} 
Co-citation, the frequency with which a pair of articles is cited by other articles,  first described independently by Small and Marshakova in 1973~\cite{small_co-citation_1973,marshakova-shaikevich_co-citation_1973},
provides insight into the emergence of new ideas derived from the association of previously independent ones. 
Co-citation thus provides some degree of evolutionary insight across disciplines and sub-disciplines and has been used for clustering publications~\cite{boyack_cocitation_2010,boyack_improving_2013,small_structure_1974,small_clustering_1985}. 
For an alternate view of these DBLP data we constructed clusters using co-citation, using a modification of   variable level clustering to the 2,685,356 articles in \emph{comp} and then performing top-down agglomerative clustering on the resultant 10,298 (of 22,232 clusters) containing greater than two nodes per cluster.

@@

Agglomerative clustering was then performed on these 10,298 clusters to generate higher-order clusters. Briefly, each cluster was now treated as a node and the edge weight between two clusters was assigned to the maximum edge weight of all edges between the nodes in these two clusters. Edges were arranged in descending order. The first pair of clusters were merged and its edge weight with other interacting clusters was recalculated. All edges were then re-ordered as before and the next pair of clusters was merged. The process was halted after 600 rounds to prevent single large clusters being generated (Fig.~\ref{fig:merge_select}).

>>>>

Other classification systems are available such as the All Science Journal Classification (ASJC) developed and maintained by Scopus, the Web of Science research and categorical classifications from Clarivate Analytics, and the NSF classification system~\cite{nsf_classification}. Of these, Scopus and the Web of Science also include global bibliographies with citation links and proprietary unique identifiers for publications, some of which can be matched to digital object identifiers (DOIs). All three rely on applying one or more journal-derived labels to articles. The inheritance, by articles, of labels meant to describe journals can be problematic in fields that evolve faster than their journal classifications. Shu and colleagues recently noted in a comparative study of the Chinese Science Citation Database (CSCD) and the Web of Science, that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. Others have also discussed and critiqued disciplinary assignments using journal based classifications~\cite{wang_large-scale_2016} \textcolor{blue}{need comprehensive list of citations}. Article classification systems have been constructed, however, which escape some of the criticisms of journal-based classification~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}.

@@

\subsection{Data}
 A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the XML source file and loaded into a PostgreSQL database. We have previously parsed the Scopus dataset of over 72,000,000 publications into a custom schema as part of implementing a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the DBLP dataset were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceedings. References cited by these publications were then extracted from Scopus (7,129,006 records), resulting in a total of 8,000,411 nodes (publications and references) and 44,296,381 edges that represent citations within the dataset. This dataset is referred to as \emph{comp} (see Fig~\ref{fig:ar_cp_annotation}).

@@

\paragraph{Clustering by Direct Citation.} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus, which requires the number of clusters to be formed as an input parameter. In preliminary experiments, we varied the number of clusters to be formed between 10 and 50 clusters (data not shown).  At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}.  Consequently, we generated 20 clusters, labeled 0-19, using Graclus (Fig~\ref{fig:heatmap}).
We used conductance, as defined in Shun et al.~\cite{shun_parallel_2016}, to  evaluate clustering (smaller is better), noting that
conductance has been found to be ``the stand-alone quality metric that best indicates performance on the information recovery metrics" \cite{emmons2016analysis}.
% The quality of clustering was evaluated by measuring conductance as defined in Shun et al.~\cite{shun_parallel_2016}. 
%George, Tandy - add text about why conductance is good, and that small values are better
The data in \emph{comp} were clustered into 18, 20, or 22 clusters and in each case the highest numbered cluster (17, 19, or 21) had the greatest conductance value and the smallest number of nodes, suggesting that it effectively served as a container for `left over publications' during the Graclus process. Thus, we did not consider cluster 19 (the 20th cluster) when interpreting our results.  The remaining 19 clusters had conductance values ranging from 0.09 to 0.25 with a median conductance of 0.15.

@@

\paragraph{Clustering by Co-Citation.}
Co-citation, the frequency with which a pair of articles is cited by other articles,  first described independently by Small and Marshakova in 1973~\cite{small_co-citation_1973,marshakova-shaikevich_co-citation_1973},
provides insight into the emergence of new ideas derived from the association of previously independent ones. 
Co-citation thus provides some degree of evolutionary insight across disciplines and sub-disciplines and has been used for clustering publications~\cite{boyack_cocitation_2010,boyack_improving_2013,small_structure_1974,small_clustering_1985}. 
For an alternate view of these DBLP data we constructed clusters using co-citation, using a modification of   variable level clustering to the 2,685,356 articles in \emph{comp} and then performing top-down agglomerative clustering on the resultant 10,298 (of 22,232 clusters) containing greater than two nodes per cluster.

@@

Agglomerative clustering was then performed on these 10,298 clusters to generate higher-order clusters. Briefly, each cluster was now treated as a node and the edge weight between two clusters was assigned to the maximum edge weight of all edges between the nodes in these two clusters. Edges were arranged in descending order. The first pair of clusters was merged and its edge weight with other interacting clusters was recalculated. All edges were then re-ordered as before and the next pair of clusters was merged. The process was halted after 600 rounds to prevent single large clusters being generated (Fig.~\ref{fig:merge_select}).
